% Put somewhere that number can be regarded as a free
% creation of the mind---if only time could be regarded so.


% Ending line should be something about how the account suggests
% that our counting is time, not simply of time.




\documentclass[12pt]{article}
\usepackage[xspace]{ellipsis}
\title{Making Time Count}
\author{Daniel Harrell}

\begin{document}
\maketitle

If you asked me the question ``What is the time?'' I could
look at my watch---or these days, at my phone---and tell
you. It's [current time=A]. But what if you changed
your question, by leaving out the definite article? Suppose
your question weren't ``What is \emph{the} time?'', but more
simply, ``What is~\dots\ time?''   

In that case, it's hard to know what my answer would be, but
easy to suppose that it could not be as immediate, or exact,
or conclusive as the time [A] was to your
first question.  And about a matter less suited to our
phones than to some philosophy about the
nature of time.

The Freshmen, in fact, have just finished reading
one---Aristotle's; and my debts to it in this lecture should
be clear before I'm done. But so will my debts to Euclid.
And the question that guides me tonight will not be ``What
is time?''---at least if asking this question means turning
away, either at once or for good, from our phones, our
watches, our clocks; or more generally, from the numbers we
use to tell time, and as if our telling of time had nothing
to tell us---to teach us---about the nature of time.  My
question, then, is something less like ``What is time?'',
and something more like: ``What is it about time that allows
it to be told?'' What is it about time that allows us to ask
about the time---definite article reintroduced---and to
answer that question with all the immediacy, the exactness,
and the conclusiveness of ``it's [A]''---or rather,
``[current time=B].'' My conjecture for the next hour, then,
will be that in telling you it's [B], or indeed in
referring to ``the next hour,'' and more generally in every
ready-at-hand way we have through numbers of marking time,
measuring time, reckoning time, there is something important
to discover, even essential, about the nature of time.
Something that accounts for time, in our counting of time.

But what could this be? After all, there is nothing
essential in my calling the next hour ``an hour.'' For I
could just as well have called this hour ``60 minutes.'' Or
even ``3,600 seconds''---if I've done my math right; and
despite the fact that the number 3,600 is not the number 60,
nor is 60 the number 1---if there is such a number. Our
counting of time in this sense, the sense in which we use
numbers to mark \emph{intervals} of time, seems borrowed and
arbitrary rather than rooted or necessary---a matter of
convention rather than essence. 

And there is even more convention, it would seem, in the
counting of time that led me to say it was [A], or [B], or,
right now, [current time=C].  This is the count where we use
numbers to mark \emph{moments} of time rather than
intervals.  For while it may be [C] in Annapolis on this
side of King George Street, on the other side, it is [twenty
[minutes of C]]---and in Santa Fe, only [six [minutes of
C]=D]. I could also call it [D] in Flagstaff---if I were
giving this lecture before Spring Break. But in giving it
tonight, I will have to admit that it is only [five [minutes
of C]=E] in Flagstaff, since Daylight Saving Time goes
unobserved in Arizona. And if I were giving this lecture
before the Uniform Time Act of 1966, I wouldn't know what to
say about the time elsewhere.  For back then, cities in the
United States could choose their own dates for Daylight
Saving Time at will; and they did, in what Time Magazine
called a ``chaos of clocks.'' Not the clocks we find
populating Einstein's theory of relativity, as if cities had
discovered themselves in different frames of reference with
no present moment in common; but rather the clocks we more
innocently use in everyday life, where there is taken to be
a present moment in common, but no like moment in number, it
seems, no ``now'' that belongs to number to stop us from
assigning the present moment different numbers at
once---where ``now'' is [current time=F] in Annapolis,
[seven [minutes of F]] in Memphis, [six [minutes of F]] in
Santa Fe, or [five [minutes of F] in Flagstaff.

Giving content to this present moment, moreover, still gives
us no sense of the number that might belong to it. The
``now'' that begins Friday Night lecture in Annapolis, for
example, is officially 8, but it used to be 8:15. In
Santa Fe it is 7:30, but used to be 8. Similarly with the
``now'' that begins seminar: for undergraduates in
Annapolis, this is 8, but for GI students
7:30---and 7:30 in Santa Fe for both.  ---Except for the
4:30 seminars, one for each class, the year I was
in Santa Fe. And if we include seminar-parties, where the
start-time is assigned at will, well---there would seem no
number excluded in principle: which is again as if there
is no ``now'' in number.  And if time itself exists
only ``now,'' in the present moment, since the past is no
longer and the future not yet, then there is nothing of
time in number.   

But then we might have inferred this much---that time is one
thing and number something else---even without such
examples. 
% Dedekind, whom the Juniors are about to read, once
% called numbers ``free creations of the human mind.'' If only
% this were true as well about the moments of time.
For if we think of time as having any existence
beyond the present moment, it is most likely when we think
of this moment as ever-new, making time 
like a river, flowing ever onward.   
% Later say time is like a march, not river.
But our counting of time uses number to form a different kind
of succession, which is never continuous, and
usually recurring: going repeatedly back to 1
from day to day, week to week, month to month, and year to
year, even as time itself rolls on---a fact soon enough
written on our faces, even if we erase it from our clocks
and calendars. 

Our counting of time in these respects even suggests that
the count is not really of time, but rather of place, in
following motions that, unlike time, continue and finally
complete themselves with respect to place. We count to 365,
then, in following the yearly trip of the earth around the
sun; as we count to 30---or close enough---in following the
monthly trip of the moon around the earth; or to 24, in
following the daily trip of the earth around itself. And
dividing this 24 into its units, accordingly divides the
earth into twenty-four places---each a so-called time zone,
explaining the difference in time between Annapolis and
Santa Fe as really a difference in place.  Even our weekly
count to 7 in this sense, which otherwise has no source in
the completed motion of any celestial body, might still be
said to follow one---the very motion that, on one account at
least, created every celestial body, along with everything
else, over six days of work finished by a seventh of rest.
Nor need we look so far above us for seeming proof that our
telling of time is really of place. For we can find this
proof writ small on the face of any watch, where every
number has its place, and every time in turn, its position.
Yet if time itself has no place or position---and we might
even consider it a negation of any place or position---then
our every clock, it seems, tells a (bald-faced) lie.

But this is to argue against what I earlier called my
conjecture for the next hour---now my conjecture for the
next [remaining time of hour=60-minutes of current time]
minutes---which is that our every clock tells the truth. And
perhaps it is time---whatever that may be---to make the case
for clocks.  Consider again our sense of time itself, when
we think of it like a river. In certain respects---decisive
respects, I think---the simile falls short. For the motion
of a river, we can see and say, is made by the river, as it
moves along its banks, now more swiftly, now more slowly.
But in the motion of time, there is nothing like the river
or its banks to see, nothing visibly moving relative to
anything else.  And no surprise, for there would only be
anything visible, it seems, if the motion of time were a
motion in space.  This motion is already peculiar, then, in
belonging to time and time alone.  And it becomes still more
peculiar when we consider its progress.  We sometimes fancy
that, again like a river moving swifter or slower, time
itself can fly or drag. But if time could really do this,
how would we know?  When anything else changes its speed,
after all, we can use some unit of time to confirm it, as a
change in the distance traveled, for example, per minute.
But we are deprived of this unit to confirm any change of
speed in time itself; for the unit is now in question: is
this minute, when compared to the last, or the next, faster
or slower, longer or shorter?  And even before this question
we must ask another, since to say that time can change its
speed assumes that time has a speed. But what could we mean
by time's speed? In what sense could time itself travel any
distance, if we are then deprived of any measure in space to
account for it? And how long would it take time to travel
this distance, if we are then deprived of any measure in
time to account for it? 

It begins to look as if our attributing any motion to time
involves a category mistake, claiming for time itself, and
apart from space, what can only happen \emph{in} time and
space together. And this correction suggests that time
itself is not a motion, but rather a medium for motion---a
qualification implicit in our sense that when something
moves from place to place, it does so \emph{in}, or even
\emph{through}, some corresponding length of time. But to
admit even this much risks another conflation of time and
space; for what could we mean by a ``length'' of time?
Indeed, the only conception of time so far that promises to
escape any such conflation with space, is one that reduces
the length of time to ``now,'' the present moment,
since the present moment can be conceived to have
no length. It makes sense to think that this
moment lasts only, well, a moment, or instant, before a new
such moment takes its place, so to speak.  Or, to borrow an
earlier formulation, it makes sense to think of the present
moment as ever-new. But this is also what tempts us to
compare time to a river. And if that analogy is false in
being spatial, then what is time really like in being
temporal?

It is just here, I think, that our counting of time deserves
some credit. For let us suppose that the present moment indeed
has no length. If it has no length, then it is no larger, we
could say, than a geometric point.  But saying this risks
another spatial analogy. For even though the point has no
part according to Euclid's definition, its integrity in this
sense, being spatial, makes the point a position, or
location. The point is where a line begins and ends, for
example; or where any two lines intersect. But the integrity
of the present moment, being temporal, is opposed to this;
and we might even characterize it as if it negated the
integrity of the point, by saying that the present moment
can occupy \emph{every} location in space, not just the one
location occupied by a point. But saying this still relies
on an appeal to space---all of space in fact. So what should
we say instead? A more promising characterization, I think,
is one I just mentioned a moment ago: namely, that the present
moment is ever-new. For as the present moment
begins, it ends, in having no parts; and as it ends,
it is replaced, in having no position. This is why the
integrity of the present moment is temporal rather than
spatial; for in having neither part nor position, this
moment establishes time as a dimension of succession rather than
simultaneity. But this gives us a way to distinguish the
present moment from the geometric point absolutely. For if
the present moment is indeed ever-new in lacking parts and
position, then one such moment will be replaced by the next
immediately, without any space between the two. And the lack of
space would be another mark of the moment's
temporal integrity. But space is precisely what
remains as a mark of spatial integrity, as the position
without parts possessed by the point, allows it to occupy
space without consuming it.  Thus, there is no way to
encounter points in succession, from one to the next;
between any two, no matter how close, there will be space
enough for a third. But the present moment, then, is not only
\emph{not} a point, it is not even \emph{like} a point.  And
if we have any hope of finding something that the present
moment is truly like, it will have to be what has neither
parts nor position.

But saying ``what has neither parts nor position,'' as it
happens, is one way to define the unit, out of which, at
least according to classical arithmetic, every number is
made.  I say ``classical,'' since we are tempted, by later
developments in the theory of number, to think of the unit
itself as a number. I even mentioned the number 1 near the
beginning of my lecture, before catching myself. And if we
do think of the unit as a number, then it is easy to suppose
it has \emph{both} position \emph{and} parts, being the
first of the so-called counting numbers, which we can always
divide by a larger such number to yield some fraction of
itself.  But it is precisely those suppositions that the
classical understanding of number calls into question. And
it is the unit on \emph{that} understanding, I now propose,
that not only makes most sense of how we count when counting
time; but is also what time is most like when we erase any
hint of space from our sense of it. Or to put this proposal
another way: we are all Euclideans when we tell the
time---and time itself is Euclidean most of all.  

To make sense of this proposal, let me first offer a simple
sketch of the likeness between number and time that I have
in mind. I just concluded that the present moment, in
lacking parts and position, is ever-new. But then we could
also say this of the unit, since it too lacks parts and
position. Or to put this another way, since the unit lacks
parts and position, we can think of it as a recurring whole,
which corresponds to the ever-new present moment.  And if we
then assign the numeral ``1'' to the unit, we can conceive
its recurrence as producing a count of the ever-new present
moment, like so: 1, 1, 1, and so on. But if the unit remains
whole in its recurrence, then it would seem as if the unit
must somehow encompass its recurrence.  That is, the unit
must somehow encompass the fact that it is being taken once,
twice, thrice~\ldots, in the count 1, 1, 1\ldots. So this
count is better conceived as starting at 1 but then
proceeding through every number in turn. So the count would
go: 1, 2, 3, and so on. And this count, now through
every number in turn, is again of the ever-new present moment, as
this moment so establishes the succession we call time. So
we can understand the count 1, 2, 3, and so on, as belonging to
time, or as being a counting of time. 

In light of this sketch, which all but identifies time with
number, my earlier reasons for denying any connection can be
reconsidered.  The first of these reasons was that there seems
to be nothing essential in my calling the next hour ``an
hour.'' For I could just as well call this hour 60 minutes,
or even 3,600 seconds. If my sketch above is correct,
however, I think we will have to say that this is mistaken.
For counting to 3,600 is different from counting to
60, while counting to 1 isn't even possible, since 1 defines
the start rather than the end of any count. But why, then,
would we think that the different numbers in this case are
equal, and equal in turn to the unit? I think it's because
we are once again conflating time and space. We refer to the
\emph{next} hour and divide it at once into seconds, as if
the hour we tell time by \emph{had} position and parts. But
when we give the unit position and parts, we are no longer
telling time by it; that is, we are no longer using the unit
for counting.  We are rather using the unit for measuring.
Or to put it another way, the unit that can indeed possess
position and parts makes the numbers we read off a ruler,
not the numbers we read off a clock.

What I mean by this is that the succession of numbers on a
ruler, say from 1 to 12, is meant to carry us through space
rather than time, in an act of measurement rather than
counting.  For when we apply the ruler to a length, the
length, being spatial, will carry us through the successive
numbers on the ruler at once. And if the length should end
somewhere between two such numbers---say between 8 and
9---there is nothing to stop us from dividing the unit
between them into as many parts as needed, again
at once, to give a more accurate reckoning of the length.
And this reckoning, so-called, will thus be a matter of
inspection rather than calculation.  But on a clock, at
least if we are using it for a reckoning of time, numbers
work radically differently, even if they look the same. And
this is why the clock tells the truth. As an example, take
the current time on my phone: [current time=G] At first
glance, this time may look like the number in my ruler
example: a number between 8 and 9, which is taken by
dividing the unit between 8 and 9 into parts---in this case,
sixty parts, or minutes.  But one sign this first
glance is deceptive is the punctuation mark I see on my
phone---a colon---that separates the 8 from the [minute of
G] And if I confirmed the time on my phone by looking at a
watch, my having to read the 8 from the position of one hand
and the [minute of G] from the position of another hand
would reflect the truth. The time [G] is really
composed of two numbers, neither of which is
reached at once, but rather in separate 
counts.  The one number is reached in a count to 8; while
the other number is reached in a count to [minute of G],
preceded by at least 8 counts to 60.  And we are stopped,
then, from dividing the unit between 8 and 9 into parts in
this case---our so-called ``next hour''---since there is no
such unit to divide: the count that reaches 8, in being a
count, hasn't yet reached 9; while the count that reaches
[minute of G], in being a count, started at 1, in a
reflection of the unit's integrity rather than divisibility.
So while different numbers may make no difference to a
measurement of time; they make all the difference to a
reckoning of time. 

In this analysis of the numbers on a clock, however, there
is a strange implication for the present moment, which I
will only mention here before returning to it later.  If we
are making two counts when we tell the time, in this case
having reached 8 by the one count and [minute of G] by the
other, then it looks as if there are two present moments:
the one that belongs to 8, and that we call ``the hour'';
and then the one that belongs to [minute of G], and that we
call ``the minute.'' It would be easy to say that the minute
is contained in the hour, reducing the two moments to one,
if we could say that the hour had parts. But I just argued
against this. Yet how could the hour be one present moment
and the minute be yet another present moment?  Especially if
the present moment, in having no parts, has no length, and
ends as soon as it begins?  Doesn't this imply that there
can be merely one such moment at once, and indeed that our
reckoning of it, whether by hours or by minutes, is rough?
Inexact? 

As I said, I will return to this matter later, but I think I
must accept the strange implication. If time and number are
identified in the way I sketched above, then I think we will
have to admit more than one present moment at once---indeed,
as many such moments as are reached in different counts. And
when I declare the time to be [G], or rather [current
time=H], I am counting to different present moments at once,
the one that still belongs to 8, and the one that now
belongs to [minute of H].  We will also have to admit, I
think, that there is nothing rough or inexact in either count, since
the present moment, in having no parts, is no bigger or
smaller than the unit in either count, which also has no
parts.  

But this brings me to the second reason I gave earlier for
thinking that time is one thing and number something else.
For if it is now [H] in Annapolis, then it is now [six
[minute of H]] in Santa Fe. Before, I took this to mean
that while Santa Fe and Annapolis share the same present
moment, they assign it different numbers, and that this
reflected a difference in place rather than time. But if my
sketch above is correct, I think we will have to say that
Santa Fe and Annapolis do \emph{not} share the same present
moment, and for reasons unrelated to any theory of
relativity. For Annapolis has reached the number 8 in a
count by hour that Santa Fe has only made to 6, which in
turn means that our respective counts by minute differ by
factors of 8 and 6. Thus, Santa Fe's present moment is
indeed different from ours, and in fact a present moment that lies
in our past, in just the way that 6 is different from 8, and
in the count to 8, already passed. So because Santa Fe is at
a lesser number in its counting of time than Annapolis,
Santa Fe is, right now, at an earlier moment in time
than Annapolis, right now, apart from any difference in place.  And
again, in a matter I will have to return to, it looks as if
identifying time with number means admitting that there can
be more than one present moment at once; and in this case we
have at least four: the two that belong to the counts of
hour and minute in Annapolis; and the two that belong to the
counts of hour and minute in Santa Fe. 

This multiplication of present moments, moreover, is
implicit in the way I would now want to correct my earlier
conclusion about the difference between time and number.
Then, I said there was no ``now''  belonging to number to
stop us from assigning the present moment different numbers
at once. But in identifying time with number, I think I will
have to say just the opposite: for the ``now'' belongs to every
number in turn; which means there is indeed something
stopping us from assigning the same present moment different
numbers at once. For if the numbers are different as the
result of different counts, then the present moments are
different as well.

This brings me to my last reason, which was more general,
for denying any connection between time and number. I said
that the present moment, being ever-new, makes time like a
river, flowing ever onward; while our counting of time uses
number to form a different kind of succession, which is
never continuous and usually recurring.  But since then, I
think I've given us better reasons for denying that time is
like a river. And on the matter of continuity, I think we
can say that temporal sucession is no different from
numerical succession, at least if we conceive the numerical
in this case, again, along classical lines. For then there
will be no gaps in our count, as every number is composed by
the unit, and reached by the unit in turn.  This is
continuity enough to reckon with the like succession of
time, where every present moment leads directly to the next,
leaving no gap in the event. 

But what about the matter of recurrence? For this
\emph{does} seem to distinguish a temporal succession, where
every moment leads directly to the next, from any numerical
succesion where we reach some number in our count only to
start the count over at 1; and for reasons, so it seemed,
having nothing to do with time, and everything to do with
space. We count to 365, or to 30, or to 24, so I said, in
following the circuits of celestial bodies. But in fact the
most important examples of a recurring count are not found
in our calendars and clocks, but rather in a waltz, or a line of
iambic pentameter. And in those cases, the sense of the
count's recurrence seems to lie within rather than without,
as we can hear the recurrence in sound without needing to
see anything in space make a circuit, or even have what we
hear in sound be made explicitly numerical. If I simply go:
da-dah, da-dah, da-dah; or, dah-da-da, dah-da-da,
dah-da-da---the recurrence of a count is plain. And this
recurrence is made articulate in the arts of poetry and
music, which we credit for having somehow shaped time rather
than space into a whole.  So even if the recurring counts in
our calendars and clocks typically track some completion in
space, these same recurring counts ought to make a temporal
sense, as if they produced a completion in time.   

But what kind of sense could this be? Indeed, how could
there be a whole in time, if the succession of moments that
constitutes time can only compose this whole part by part,
but leaves every such part at \emph{that} moment in time,
preventing the parts from ever coming together as a whole? I
think there is an answer to this question; and it returns
me, at last, to the strange implication about the present
moment that I mentioned before. For let us suppose there
\emph{can} be more than one present moment at once. As an
example, take again the time on my phone, which is now
[current time=I]. And yet again, I will intepret this time
as indicating two present moments, the hour that belongs to
8, and the minute that belongs to [minute of I]. And to
repeat my earlier point about this: it would be easy to
reduce the two present moments to one, if we could say that
this hour contained this minute. But we can't say this
without wrongly dividing the hour we count by into parts.
But what I think we \emph{can} say is exactly the opposite:
it is not that the hour contains the minute, but rather that
the minute contains the hour. For in belonging to the number
[minute of I], this minute already encompasses the previous
[minute of I-1] minutes by our count, as will every minute
that follows in our count---until we start the count over
again. And we start the minute-count over again, not to
follow the circuit of any celestial body, but rather to
contain and comprehend the hour---in this case, the hour
that belongs to 8---which remains present and whole
throughout the sucession of present moments that we count by
minute from 1 to 60. If this analysis is right, then it
implies that there could never be genuine temporal wholes,
along with any truth in the poetry and music that achieve
them, unless there could be more than one present moment at
once. And anytime we encounter such a whole, we are
experiencing at least two present moments at once,
corresponding to two counts of time, where one count recurs
in relation to the other one. 
	
There is, however, a question that remains if everything I
have said so far is correct. For again, I have all but
identified time with number by reducing them both to
elements that lack any parts or position. It is difficult,
then, to tell any difference between the present
moment and the unit; and difficult in turn to tell any
difference between time and number. But is there
any difference? This question, I think, is what
Aristotle's account of time just read by the Freshman is
intended to answer; and I hope my lecture has given some
indication of how radical an answer it might be. For
I read his answer to be simply no.  There is no difference between time and
number, because time \emph{is} number. And if Aristotle is
right, then our telling of time tells the truth. And the
next time someone asks you ``but what, then, is time?'' you
can look at your phone, and say: it's [current time=J].

\newpage
\appendix
\section{Time Sheet}

\noindent 

Time A=\\

\bigskip

Time B=\\

\bigskip

Time C=\\

\bigskip

Time D=\\

\bigskip

Time E=\\

\bigskip

Time F=\\

\bigskip

Time G=\\

\bigskip

Time H=\\

\bigskip

Time I=\\

\bigskip

Time J
\end{document}
% Note for later: the cumulativeness of number,
% corresponding to time, is why there can be different
% counts.
a difference at the stroke of midnight in Annapolis, for
example, as to whether a day, or month, or year has yet the
new year as to whether the birthday of a friend in Santa Fe
has yet arrived, But since I'm about to conclude things
about the present moment even stranger than this
duplication, I will leave my defense of this first
conclusion at this: since it's only [current time - 2 hrs]
in Santa Fe right now, isn't it indeed earlier in Santa Fe
right now? 

han in Annapolis right now?  y present moment is different
from Santa Fe's, and later than Santa Fe's, in just the way
that 8 is different from 6, being later in the count?  in
Santa Fe, 
% Move on to present moments different; allude to later
% discussion of present moment, which is then discussed in
% discreteness and recurrence discussion. Time is
% cumulative, You can be in one present moment at the same
% time. Also talk about the wholeness of the present moment,
% this is what makes it cumulative.  explain cumulativeness
% by reducing a succession of unit, unit, etc to repetition
% without displacement. Difference between being in a
% succession and producing one: 1, 1, 1, doesn't capture the
% difference of ordinal position. Or, the wholeness of the
% unit can't be negated by its repeated displacement. What
% Hegel calls determinate negation. Present moment has
% depth. End
% with: counting can only belong to time, or be of time, if
% it is time.  Time not divided, but differentiated, by a
% present moment within a present moment. Also, present
% moment is whole.  Identify this both with its variability
% and the way it accumlulates End with, the counting we do
% of time, is time.

%Recurrence is the temporal analogue to division.

% Perhaps mention issue of whether time is finally
% terrestrial or celestial, cyclical or linear. 
% Recurrence Not to comprehend a place in space, but rather a place in
% time.

involves no division of any unit, in being a count that
started at 1, 

[Now go on to explain how this makes the unit disappear,
using past formulations, that there is something to stop
us.]

In
neither count to the first is a count to 8, while the second
a count to whether the unit is an hour; while the second is
a count to more coand the second, the colon I see between
the 8 and the [seconds of current time] is a sign that this
first glance is deceptive; and if we read the same so-called
number off a watch, dividing the first number requiring us
to look at a separate hour and minute hand, with an hour and
minute hand, the truth is easier to see. is also a number
between 8 and 9, but we have ; and at first glance, it looks
as if a unit between 8 and 9 has been divided
into as many parts as needed to give a more accurate
reckoning of the time. But we have not reached this point
ell the time more give he unit along with the numbers it
composes has to be revised, radically, But what happens,
then, when we use the unit for counting rather than
measuring? And how does this unit compose the numbers we
read off a clock? Well, let tell the time? 

% Relativity of present moment present moment variable, not
% in size, but rather in the scope of the count,
% corresponding to the scope of your horizons. If you are
% prepared to make a longer count, then the horizons of your
% present moment will shrink. Or if you want to relax and
% let the count be shorter, your present moment will expand.
% rather than divisible. If you're prepared to count to 60,
% for example, then the present moment will be a minute. to 12,
% the present moment is a month, if you're prepared to count
% to 30, the present moment is  
for
we can simply divide the unit-distance between them
into as many parts as needed, again at once, to measureit
becomes intelligible to suppose that the unit some mark on
the ruler at once, a corresponding number on the ruler at
once. --suppose the ruler is marked in inches it to be 6
inches---the length will carry us to the number 6 on the
ruler at once; and also at once to any number equal to 6
according to some subdivision of the ruler.  a corresponding
any some subdivision at once, as well as at once to any
other number, say 12, that is equal to 6 according to some
subdivision ; and it will also be at once that we can
subdivide the interval between 5 and 6 ; and  it falls a bit
short of the number 6, we can we canThis is why we can read
any number off a ruler at once, for a matching length, the
numbers off a ruler at once, by inspection; as well as why
we can read off divisions between successive numbers
immediately. if the length we measure reaches the number 6,
we can simply look at the ruler , without needing to count
One sign of this is that we can read the numbers off a ruler
by inspection; while This is why we can read the number off
a ruler

% So in another sense I was incorrect saying I could just as
% well call the next hour the next 60 minutes. 
To begin to see this, 
was I actually correct in saying this?
In one sense, I think I was. And this is the sense in which
we can indeed refer to the next hour as if the hour had
position, and divide this hour into seconds as if the hour
had parts.  Or more generally, I was correct in the sense
that we can indeed attribute both position and parts to the
unit by which we tell time. But when we give the unit parts
and position like this, I want to suggest, we are no longer
using the unit to count; we are rather using the unit to
measure. In other words, this is the unit that makes the
numbers we read off a ruler rather than a clock.But so what?
Well, for one thing, to talk of the ``space'' between
numbers being ``filled'' by the unit risks(Mention problem
with space here, then clarify by example showing that we
think of unit as having position, arranging numbers on a
number line.) Well, to see what, consider the count from 8
to 9, which numbers correspond, more or less, to the time of
the lecture I'm giving right now, from 8:00 to 9:00, taking
the unit of the count in that case to be the hour.  On my
proposal, this unit should be understood as having no parts;
therefore, it cannot be divided. But surely it \emph{can} be
divided.  Otherwise, what could it mean to say, quite
intelligibly, that I have been at this podium already for
roughly [current time-8:00 minutes] of this hour? Or that I
will be standing here for roughly [9:00-current time
minutes] more? Or to look at my phone and declare that it is
exactly [current time]? In all these cases, I seem to be
dividing the unit that allows me to count by hour from 8 to
9. And that the unit in this case is obviously, almost
self-evidently divisible, is what causes us to think, as I
noted at the beginning of my lecture, that there is nothing
essential in my calling the next hour an hour. For I could
just as well have called it 60 minutes, or even 3,600
seconds.  

But there is something to notice in all this division that
is easy to overlook. For when we divide the unit like this
meaningfully, it is to give the hour not simply parts, but
position. I think of the next hour as
having parts like this, I am also giving it a position,
which is why I can call it the ``next'' hour. Similarly with
the present hour, when I think of it as \emph{this} hour, to
distinguish it from the last or next.  And our division of
the hour is most often done to establish this position more
precisely.  (Speak of where I am in the current lecture.) So
when we give the unit parts, we are also giving it position.
cyySimilarly when I think of the present hour as having
parts--Or if I think of the present hour as having partsor
the present hour as h like this, there is next hour ahead of
me that I can divide into minutes or seconds, or when I
think there is a present hour as divisible into minutes or
seconds, or when I think I am right now at some particular
place in the current hour, I now want to suggest that our
intuitions about this are wrong. 

% So in another sense I was incorrect saying I could just as
% well call the next hour the next 60 minutes. 

(Now go on to write about how the unit has integrity like
temporal integrity in counting; while the unit has spatial
integrity in measuring. Like we conflate space and time, we
conflate rulers with clocks, and the way we use numbers on
both. Go into this difference, with the purpose being to
show how the unit is divisible as a matter of measurement,
but not as a matter of counting. Compare ruler-face with
watch-face. For example, while 60 minutes might be equal to
3600 seconds, involving a unit subdivided (1 minute=60
seconds), once we speak of these numbers temporally, in
terms of a count, then we are counting from 1 to 60 to get
to the minute. the unit remains whole in the count. Then go
back to abstract discussion, show that if we treat the unit
spatially, then we don't correct for the difference between
points and present moments; there seem to be units in
"locations" in each number, and a gap between any two
counting numbers that can be filled in. But if we take
numbers to be generated in counting, there there is only one
unit,  "now", which is then added to get any number, and
fills the gap. Also point out that saying "one" defines the
unit by which numbers are subsequently counted. This can
vary. Of no particular size.  One implication of the unit so
understood can be formulated now.  For if the present moment
were most like the unit in this sense, then we could still
grant that no number belongs to this moment intrinsically,
but only because what belongs to it intrinsically is the
unit that composes every such number.  But a second
implication is readily inferred in the contrast between what
the point cannot do and what the unit can.  For points, so
we observed, have nothing by which they might be put
together, to form a line of any length. But this is
inevitably what units do when they are joined  


% Contrast unit put together with unit, and point put
% together with point. Perhaps even talk about discreteness
% and recurrence.

% Then move on in next
% paragraph to how to really read a watch. Perhaps then move
% on to poetry and music.)
But there is a
paradox here, which suggests that time can only move forward
if it is \emph{not} like a flow. For it is not as if going
forward in time gives time any length it lacked in the
present moment, since it takes length to produce length. But
is nothing in this moment that can join it to any other such
moment, the way that two lengths can be joined on one side
at a point.  This lack of length wouldn't preclude the
present moment from being ever-new; it would only preclude
the result of its being ever-new from having continuity. Or
to put the result more positively: if the present moment
were ever-new, but lacked any length, then this would
produce a succession of such moments.
% Later deepen this conception through the distinction
% between a succession or sequence, and a series.
But this succession would carry us forward not like the flow
of a river, but rather like the count to a number, in which
the unit is also, in a sense, ever-new---as it repeatedly,
or ever-newly, carries us from 1 to 2, then from 2 to 3,
then from 3 to 4, and so on. 

[This should be a new paragraph.] What is more, because we are
now counting our way from number to number, the difference
between numbers that make its domain discrete rather than
continuous is in one sense overcome. 3 is separated from 2
by a unit, to be sure; but it is by this same unit that we
get from  2 to 3 in a count. Or to put the point a different
way, while 3 may be composed of three separate units as a
number, it is composed of three units added together as a
sum. And it suggests, if we now take counting to be what
time is most like in going forward, that what we initially
take to be the continuity in this going-forward is really an
integrity. 

this
adding-together that leads to any number in a count, would
seem to give sufficient integrity to the succession 


to be sure; but it is this very unit thwhile this succession
of numbers is no doubt discontinuous, the result of
this count is discontinous If time's likeness to counting seems
implausible so described, Moreover, it would only be true in one
sense that this count was discontinuous: 

Continuity vs integrity. 

Now move on to difference between counting and measuring.

it seems, is to rid the point of any position, so comparing
the present moment to what has neither position nor part. rid the analogy of any position, making the
present moment like what has neither position nor part. But
the only thing like this is unit. 

or the ``now'', as i(Move on to present moment.)  ,  in
truth by a length of time? The equivocal use of ``length''
here indicates the problem with (Again, the problem here is
spatial.)

they travel, let us imagine, not simply across a given length in space-   
and this is part being time., precisely because time is time. this: 
precisely because time is time, it cannot move. But then, is
it doing, These reasons are sufficient, it seems to me, to look
elsewhere We can press the problems in our sense of time itselfYet as I mentioned above, we liken time to a river, I think, in the
attempt to conceive its existence beyond the present moment. ,
which means thinking of this moment as ever-new or
-changing. i:w:[Write from here.]
Time, then, seems not to be a motion.
[Now turn to the present moment or now, and the form of
succession.]
measured.  in a motion still more
peculiar the motion of time were always regular rather than ever-varying,
but then in a motion, according to
time, in units of time, as if
it could vl,

one in which nothing is seen to move Time's
motion is then peculiar:If we then
insist that time is still a motion, it remains a peculiar
motion in which there is no discernible object moving, or
any objects again in our sense
of time's like motion, Consider again First, there is
something amiss paradoxical and even incoherent in our sense
of time itself. We might again suppose it flows like a
stream, or river, but presumably without any change of
actual place or ; yet time is surely unlike a stream or
river in more decisive respects: for this is presumably
without any change in place, If we think of it as a stream,
or flow, we are thinking of it by analogy to a river; but
the analogy, when pressed, falls apart almost at once. For
one thing, the flow of a river is a motion, involving both a
change of place and a thing doing the changing. ---let us
suppose these belong to the bank, andes a change of place
along with an identifiable sourceis a motion that involves a
discernible change of place; whereas the flow of
timepresumably involv every clockthat again there something
important and even essential to discover about the nature of
time in our, even essential, about the nature of time in our
is is the best I can do to argue against what I called my
working hypothesis
% Things to put here: Time unlike a stream in important
% ways: no change of place, but contains every change of
% place; a motion without an object; streams can go faster
% and slower, while time cannot. Also, point out paradox of
% now: if this is all time is, then time cannot be
% continuous in being composed of nows, and involves a kind
% of succession

And if our counting of time were really of place, it would
explain not simply the recurrence in our count, but the
discreteness in number that makes it a count. For then we
could say that each such number corresponds to a place in
the motion counted. And our clocks, so understood, would be
no different from our rulers: dividing a continuous
magnitude into so many successive but contiguous parts, in a
count from, say, 1 to 12. The count will be of inches if we
find this magnitude laid out on a ruler; but if we rather
find it rounded off, as on the face of watch, then the count
will be of hours. In both cases, however, the count from 1
to 12 will be following a change in place, not time.
The only difference, it seems, will
be one that underlines the sense in which our apparent
counting of time is really of place.  For the count from 1
to 12 on a ruler can take place at once, while the count
from 1 to 12 on a watch---well, we might have said it takes
time, but according to the account just given, we should
rather say it takes motion: in this case, the motion of an
hour-hand, inching its way around the watch's circumference,
which in turn follows the motion of the earth around its own
circumference. And so it goes if we count instead to 60 on
our watch, now following the faster motion of a minute-hand
or still-faster motion of a second-hand. All these cases
suggest that a change in place is not simply what our
apparent counting of time is really of, but even what it
really is if temporal at all: involving one motion, like the
rotation of the earth, being counted by another motion, like
the hour-hand of a watch. Perhaps time itself is yet a third
motion, encompassing the first two and any other; but it is
hard to see how this motion is Similarly,  Just our count
from 1 to 60 on a watch the count from 1 to  we can  

But this much comparison between clocks and rulers obscures
an essential difference. For the count from 1 to 12 on a
ruler takes place at once (perhaps put taking time up here
immediately, mention circularity, but then try to analyze it
out of the account to make sense of time) mention
simultaneity here).  And this is why, when we use a ruler to
count, we reach any number---say, 6 inches---by inspection
rather than calculation. since its parts, or places, exist
alongside one another, in a relation of simultaneity. And in
, the count from 1 to 12 on a watch---well, it takes time,
or at least takes place in a kind of succession that invites
us to re-consider the nature of this count.  the count is a
counting of. In a sense, we ourselves cannot count to 12 on
a watch, just like that; we have to wait for the hour-hand
to do it, as it inches its way round the circumference. The
waiting is less if we follow the minute-hand around the
circumference, and still less for the second-hand, but this
only underlines the point. In every such case, and it seems
to make for the difference between a temporal count and a
merely spatial count, the count in question follows the
change in place not directly, but through a motion. The
count itself is not what carries us from place to place, but  

Also mention that on a ruler, the places themselves are
numbered, while on a watch, it is only a motion through
those places that is numbered.
And this reflects a difference
in what we mean by \emph{counting} in either case. For in
the case of the ruler, we can redivide the magnitude into,
say, sixty parts, and our count to 60 will be equal to our
count to 12, by encompassing the same interval of length.
But on a watch-face, this same redivision of its
circumference, from 12 parts into 60, makes for unequal
counts, since they encompass different intervals of time. In
the case of a ruler, our count from 1 to 12 can be remade ,
we could say that while the count from 1 to 12 on a ruler
[Distinguish second, minute, hour hand, and 12 from 60;
different rates of speed.] Note too that the divisions of
the same magnitude also take different times. Also how this
underlines the sense in which we are indeed counting a time.
We have to wait for the .  Link to count in sense that the
different times each have a different count. 

it will be a
magnitude that we measure And this would be by analogy to the
way we measure any magnitude, which may be continuous in
itself but is divided into parts And just as the recurrence
of our countThis not only explains recurrence of our count,
but the discreteness of the numbers we use in the count.
Start this paragraph by now talking about how the the
continuity of time is in truth a continuity in the places of
the motion, and our use of numbers here is no different than
our use of them is measuring places with a ruler. [Put how
this is exactly like the measure of a ruler]

Mention how the count itself can be regarded as a motion,
and use Einstein, along with the sense in which larger
time-scales depend on the clock time-scale of each passing
hour, to support this. 
But then use the Aristotle to counter that. This part will
then conclude with the observation that we have altogether
separating our count of time from time. How to restore them?
Then go on to use the Aristotle.


Brief show how change explains everything about the previous
examples. Even to the point of thinking that the count tself
is a kind of change (refer to Einstein), leaving nothing to
time.  Then use Aristotle as transition.  but even that it
isn't of time that we do it.

please.  n number, that stops us
from  their clocks inevitably out-of-sync,  paper on the
special theory of relativity, as if cities had suddenly in
this case, I take it the magazine didn't mean the chaos in
this case seems to have come not from any deep truth about
time, as if cities had suddenly discovered the special
theory of relativity; but more simply because there is
nothing in the nature of a given temporal moment that
demands it be assigned a given number---or indeed any
number. Or to put it another way, there is no ``now'' in
number. demands it be assigned a given number ``now'', that
demands it be assigned a particular number. o preclude any
number.  relativity of simultaneity in different coordinate
systems of reference, presumably This chaos, presumably, was
produced not by any recognition, say, of

since there was no set date about when
a before the Uniform Time Act of
1966, since there were about just when
Daylight Saving Time could start or stop, a
particular place laimed to
have produced a ``chaos of clocks'' since there was no law
then  first conceived, it seems, by Benjamin
Franklin, to reduce the costs of candlelight.  to count the moments of time more truly, but
rather to speas a way of spending less on candles for light,
rather than as a way of  reducing the cost of candlelight
rather than of counting  
rather than of counting the to
count not to count the moments of time more truly, but
rather to reduce the cost of candlelight. 
(start from here) 

even
more conventional and arbitrary would seem And our
way of counting numeration . iesOne might
even say that philosophical reflection. Mention Augustine
and Aristotle, along with ignorance and non-existence, about
the nature of time. As if time were something mysterious and
deep. Then proceed to a lecture keeping closer
to the connection, since there seems to be one. And that
reflection can start where we are. [Now move on to why the
separation.] 


Why in that case is there a seeming distanceeven
gapbetween the ease of answering the question "what is the
time?" and the difficulty of answering the question "what is
timee?"

% Unobvious implications: present moment is relative, while
% the counting of time is absolute (thus, the now in Santa
% fe is different from the now here, by the count of the
% time, not according to relativity; it is indeed earlier
% there than here. We have already experienced the now they
% have experienced. Not simply according to the sun, but
% according to time.) Also, time is not continuous, but it
% does have integrity. Also, it does recur. Or rather: the
% recurrence is in the forward count; starting over marks an
% advance, as if to mark something new rather than starting
% over. Start with
% brief indication that only the counting numbers can be
% encountered in succession; once we introduce rationals, we
% lose this property. 

% Because the present moment has neither parts nor position,
% it exists only to be immediately replaced by the next such
% moment. At end of lecture raise question as to whether the
% passing of time is any different from the counting of
% number.
% Put this below: Or to put this proposal
% another way, by allusion to an earlier point: no number
% belongs intrinsically to the present moment, true; but only
% because what belongs to this moment intrinsically is the
% unit out of which every such number is made. 

% Start with reference back to succession: we can put the
% counting numbers in order from one to the next; if unit is like a
% point, then the difference between numbers is absolute;
% but in counting by a unit, we overcome this difference.

% when I return to discreteness and recurrence issue,
% distinguish between 3=1+1+1 and 3=2+1.Not adding one to
% one (most recent), but rather to beginning of count.
% Also use this account to show how number say something
% about time being non-continuous and recurring, in having
% integrity and being cumulative. "New day" follows last day
% as a whole, and the date goes forward one. 
% 
% recurrence, depending on how far count has gotten away
% from unit, is not a way of returning, but a way of moving
% forward. Meaning that our count is cumulative. Of rounding
% off and

